---
output: pdf_document
---
#Angle Analysis 

## Plants

-Population 1 was done from photographs on older plants, where we could only retrieve the angle for L1 - L4.

-Population 2 was performed on plants around 20 days old and was calculated from histology. 

-Population 3 was measured from plants around 1 months old from only PIN1::GFP x e2/e2

## Analysis

Libraries

```{r}
library(reshape2)
library(ggplot2)
```

Do analysis on 

```{r}

angleData <- read.csv("../data/anglePopulation1&2&3.csv")
head(angleData)
angleDataLeaf <- angleData[,c(2,3,4,5,6)]
angleDataSub <- subset(angleDataLeaf, leaf != "L7L8")
levels(angleDataSub$genotype) <- c("WT","e-2", "complemented e-2")
```

Make a boxplot that illustrates angle divergence.

```{r}
ggplot(angleDataSub, aes(leaf, angle)) + 
  geom_boxplot(outlier.size = 0) +
  geom_point(position = "jitter", alpha = .2) + 
  facet_grid(.~genotype) + 
  scale_color_brewer(palette="Dark2") + 
  theme_bw() + 
  geom_hline(yintercept=137.5, linetype = "dotted") + 
  scale_y_continuous(breaks=c(50, 100, 137.5, 150, 200)) + 
  theme(axis.title.x = element_text(face="bold", size=30), 
        axis.title.y = element_text(face="bold", size=30), 
        axis.text.x  = element_text(size=16),
        axis.text.y  = element_text(size=16),
        legend.title = element_text(size=30),
        legend.text = element_text(size=20),
        strip.text.x = element_text(size = 20),
        legend.position="none")
```

Make a diagram illustrating how many plants of e2 and wt are distichous, spiral & other.

Spiral = Both L2L3 and L3L4 between 147.5 and 127.5.
Decussate = Leaf L2L3 between 80 to 100 and L3L4 between 190 to 170.
Other = Everything else. 

At this point it may be beneficial to change to the wide format where each plant in a single observation by casting. 

```{r}
angleDataPlant <- angleData[,-c(6,1)]
castAngleData <- dcast(angleDataPlant, plant + genotype ~ leaf, value.var = "angle")
```

How many of each genotype
```{r}
summary(castAngleData$genotype) 

wtAngle <- subset(castAngleData, genotype == "3130")

#Spiral WT
wtSpiralL34 <- subset(wtAngle, L3L4 <= 150 & L3L4 >= 120)
dim(wtSpiralL34) #75

wtDecL34 <- subset(wtAngle, L3L4 <= 195 & L3L4 >= 165)
dim(wtDecL34) #5

##e2

e2Angle <- subset(castAngleData, genotype == "e2")
dim(e2Angle)

#Spiral e2
e2SpiralL34 <- subset(e2Angle, L3L4 <= 150 & L3L4 >= 120)
dim(e2SpiralL34) #18

e2DecL34 <- subset(e2Angle, L3L4 <= 195 & L3L4 >= 165)
dim(e2DecL34) #39
39/88
44+39
83/88

#How many plant were recorded for L3L4
summary(is.na(e2Angle$L3L4)) #87

#Other
88 - 39 -18 #=31
```

A few other ways to look at this data 

```{r}
ggplot(angleDataSub, aes(angle, fill = leaf)) + 
  geom_bar() +
  facet_grid(.~genotype) + 
  scale_color_brewer(palette="Dark2") + 
  theme_bw() +  
  scale_x_continuous(breaks=c(50, 90, 137.5, 180, 200)) + 
  theme(axis.title.x = element_text(face="bold", size=30), 
        axis.title.y = element_text(face="bold", size=30), 
        axis.text.x  = element_text(size=16),
        axis.text.y  = element_text(size=16),
        legend.title = element_text(size=30),
        legend.text = element_text(size=20),
        strip.text.x = element_text(size = 20),
        legend.position="none"
        )

angleDataSubL12 <- subset(angleDataSub, leaf == "L1L2")
angleDataSubL23 <- subset(angleDataSub, leaf == "L2L3")
angleDataSubL34 <- subset(angleDataSub, leaf == "L3L4")
angleDataSubL45 <- subset(angleDataSub, leaf == "L4L5")

angleDataSubSub <- rbind(angleDataSubL12, angleDataSubL23, angleDataSubL34, angleDataSubL45)

ggplot(angleDataSubSub, aes(angle)) + 
  geom_bar() +
  facet_grid(leaf~genotype) + 
  scale_color_brewer(palette="Dark2") + 
  theme_bw() +  
  scale_x_continuous(breaks=c(50, 90, 137.5, 180, 200)) + 
  theme(axis.title.x = element_text(face="bold", size=30), 
        axis.title.y = element_text(face="bold", size=30), 
        axis.text.x  = element_text(size=16),
        axis.text.y  = element_text(size=16),
        legend.title = element_text(size=30),
        legend.text = element_text(size=20),
        strip.text.x = element_text(size = 20),
        strip.text.y = element_text(size = 18),
        strip.background = element_rect(fill="#b0c4de"),
        legend.position="none"
        )



ggplot(angleDataSubSub, aes(angle)) + 
  geom_bar() +
  facet_grid(.~genotype) + 
  theme_bw() +  
  scale_x_continuous(breaks=c(50, 90, 137.5, 180, 200)) + 
  geom_vline(xintercept=137.5, color = "red", linetype= "longdash") +
  theme(axis.title.x = element_text(face="bold", size=30), 
        axis.title.y = element_text(face="bold", size=30), 
        axis.text.x  = element_text(size=16),
        axis.text.y  = element_text(size=16),
        legend.title = element_text(size=30),
        legend.text = element_text(size=20),
        strip.text.x = element_text(size = 20),
        strip.text.y = element_text(size = 18),
        strip.background = element_rect(fill="#b0c4de"),
        legend.position="none"
        )





ggplot(angleDataSub, aes(angle)) + 
  geom_density() +
  facet_grid(.~genotype) + 
  scale_color_brewer(palette="Dark2") + 
  theme_bw() +  
  scale_x_continuous(breaks=c(50, 90, 137.5, 180, 200)) + 
  theme(axis.title.x = element_text(face="bold", size=30), 
        axis.title.y = element_text(face="bold", size=30), 
        axis.text.x  = element_text(size=16),
        axis.text.y  = element_text(size=16),
        legend.title = element_text(size=30),
        legend.text = element_text(size=20),
        strip.text.x = element_text(size = 20),
        legend.position="none"
        )

ggplot(angleDataSub, aes(angle, fill = leaf, alpha = 50)) + 
  geom_density() +
  facet_grid(.~genotype) + 
  scale_color_brewer(palette="Dark2") + 
  theme_bw() +  
  scale_x_continuous(breaks=c(50, 90, 137.5, 180, 200)) + 
  theme(axis.title.x = element_text(face="bold", size=30), 
        axis.title.y = element_text(face="bold", size=30), 
        axis.text.x  = element_text(size=16),
        axis.text.y  = element_text(size=16),
        legend.title = element_text(size=30),
        legend.text = element_text(size=20),
        strip.text.x = element_text(size = 20)
        )
```


